PLANETESIMAL ACCRETION AT SHORT ORBITAL PERIODS

### 2021-11-24 main.tex
Gas drag and viscous stirring equations

In the `nominal' case, we give the bodies a bulk density of 3 g cm$^{-3}$, and $\left< e^{2} \right>^{1/2} = 4 e_{h}$, where $e_{h} = e a/r_{h}$ (define this in the previous section instead) which corresponds to $\alpha = 3.6 \times 10^{-2}$ and $\beta = 3.4 \times 10^{-3}$. These parameters are chosen to match the initial conditions of \citet{kokubo98}, which gave rise to oligarchic growth. To vary the value of $\alpha$, we alter the bulk density of the particles. In the high-$\alpha$ case, the bulk density is reduced by a factor of $\sim$ 7100, which produces $\alpha = 1$. To vary $\beta$, the eccentricity dispersion is increased. For the high-$\beta$ case, $\left< e^{2} \right>^{1/2}$ is increased to $1500 e_{h}$, which corresponds to $\beta = 15,000$.
% Should I choose beta = 1 instead? I did beta >> 1 to see if inelastic mergers will cool things off. Havent decided if that's important here
% Another concern: using too large of a beta makes the apo-peri distances larger than the width of the annulus

@@

Motivated by the apparent dependence of accretion modes on $\alpha$, we next investigate whether this highly efficient, non-oligarchic growth mode should be expected to operate in nature. Given the dearth of short period terrestrial planets observed around M stars (citation), we model a series of much wider disks, which span from 1 to 100 days in orbital period orbiting a late-type M star of mass 0.08 $M_{\odot}$. For a population of planetesimals with a bulk density of 3 g cm$^{-3}$, this orbital period range corresponds to $\alpha \in (0.7, 0.05)$. By simultaneously modeling a broad range of orbital periods, we can determine the critical value of $\alpha$ that divides these two modes of accretion, and also explore how the oligarchic/non-oligarchic accretion boundary affects the resulting distribution of protoplanets.

@@

\noindent where $M_{*}$ is the mass of the central star, $\Sigma_{MMSN} = 10$ g cm$^{-2}$ is surface density of the minimum-mass solar nebula \citep{hayashi81} at 1 AU, and $A$ is an enhancement factor. In the first case (fdHi), we model a disk that follows the surface density

@@

\section{Radial Transport of Material}

% from how far do embryos accumulate material
% do embryos migrate at all? might need to run a simulation without gas to demonstrate whether this is due to PDM

>>>>

% Mention Rayleigh distribution
In the `nominal' case, we give the bodies a bulk density of 3 g cm$^{-3}$, and $\left< e^{2} \right>^{1/2} = 4 e_{h}$, where $e_{h} = e a/r_{h}$ (define this in the previous section instead) which corresponds to $\alpha = 3.6 \times 10^{-2}$ and $\beta = 3.4 \times 10^{-3}$. These parameters are chosen to match the initial conditions of \citet{kokubo98}, which gave rise to oligarchic growth. To vary the value of $\alpha$, we alter the bulk density of the particles. In the high-$\alpha$ case, the bulk density is reduced by a factor of $\sim$ 7100, which produces $\alpha = 1$. To vary $\beta$, the eccentricity dispersion is increased. For the high-$\beta$ case, $\left< e^{2} \right>^{1/2}$ is increased to $1500 e_{h}$, which corresponds to $\beta = 15,000$.
% Important note: need to use f=6 for the low alpha case, otherwise the simulation takes forever. Do these alpha values reflect the artifically enhanced collision cross section im using?
% Should I choose beta = 1 instead? I did beta >> 1 to see if inelastic mergers will cool things off. Havent decided if that's important here
% Another concern: using too large of a beta makes the apo-peri distances larger than the width of the annulus

@@

\subsection{Initial Conditions}

Motivated by the apparent dependence of accretion modes on $\alpha$, we next investigate whether this highly efficient, non-oligarchic growth mode should be expected to operate in a typical planet-forming disk. Given the dearth of short period terrestrial planets observed around M stars (citation), we model a series of wide planetesimal disks, which span from 1 to 100 days in orbital period orbiting a late-type M star of mass 0.08 $M_{\odot}$. For a population of planetesimals with a bulk density of 3 g cm$^{-3}$, this orbital period range corresponds to $\alpha \in (0.7, 0.05)$. By simultaneously modeling a broad range of orbital periods, we can determine the critical value of $\alpha$ that divides these two modes of accretion, and also explore how the oligarchic/non-oligarchic accretion boundary affects the resulting distribution of protoplanets.

@@

\noindent where $M_{*}$ is the mass of the central star, $\Sigma_{MMSN} = 10$ g cm$^{-2}$ is surface density of the minimum-mass solar nebula \citep{hayashi81} at 1 AU, and $A$ is an enhancement factor. In the first case (fdHi), we model a disk that follows a MMSN power law slope, with the overall normalization enhanced by a factor of 100. This choice of parameters for the solid surface density profile appears necessary in order to reproduce many observed short period terrestrial worlds in-situ \citep{hansen12}. Additionally, we vary the power law index (fdHiShallow, fdHiSteep) and overall normalization (fdLo) of the solids profile.

In all cases, the eccentricities and inclinations of the bodies are randomly drawn from a Rayleigh distribution, with $\left< e^{2} \right>^{1/2} = 2\left<i^{2} \right>^{1/2} = e_{eq}$. The value of $e_{eq}$ is chosen such that the timescales for viscous stirring and aerodynamic gas drag on the planetesimals are in equilibrium. The viscous stirring timescale is given by \citet{ida93} as

\begin{equation}\label{eq:vs_timescale}
    \tau_{vs}  = \frac{\left< e^2 \right>}{d \left< e^2 \right> / dt} \approx \frac{1}{40}\left(\frac{\Omega^{2} a^{3}}{2 G m_{pl}}\right)^{2} \frac{4 m_{pl} \langle e^{2} \rangle^{2}}{\Sigma a^{2} \Omega},
\end{equation}

\noindent where $\Omega$, $a$ and $e$ are the orbital frequencies, semimajor axes and eccentricities of the individual planetesimals. In the Stokes regime, the gas drag timescale is given by \citet{adachi76} as

\begin{equation}\label{eq:ts_stokes}
    t_{s} = \frac{2 m_{pl}}{C_{D} \pi r_{pl}^{2} \rho_{g} v_{g}},
\end{equation}

\noindent where $C_{D}$ is a drag coefficient of order unity, $\rho_{g}$ is the local gas volume density and $v_{g}$ is the headwind velocity of the gas experienced by the planetesimal. The local gas volume density is given by

\begin{equation}\label{eq:rho_gas}
	\rho_{g} = \frac{\Sigma_{g}}{\sqrt{2 \pi} h_{g} \exp\left[ -z^{2} / \left( 2 h_{g}^{2} \right) \right]},
\end{equation}

\noindent where $\Sigma_{g}$ is the gas surface density (taken to be 240x the solid surface density), $h_{g} = c_{s} / \Omega$ is the local gas scale height and $z$ is the height above the disk midplane. The sound speed profile is given by $c_{s} = \sqrt{k_{B} T(r) / \left( \mu m_{H} \right)}$, where $k_{B}$ is Boltzmann's constant, $T(r) = T_{0} r^{-Q}$, $\mu = 2.34$ and $m_{h}$ is the mass of a hydrogen atom. For a protoplanetary disk around a typical M star, $T_{0} = 148$ K and $Q$ = 0.58 \citep{andrews05}.

Finally, the headwind velocity of the gas, due to the fact that the gas disk is pressure supported, is given by

\begin{equation}\label{eq:v_gas}
	v_{g} = v_{k} \left[ 1 - \sqrt{ Q c_{s}^2 / v_{k}^2} \right],
\end{equation}

\noindent where $v_{k}$ is the local Keplerian velocity.

@@

\section{Assembly History of Embryos}

% from how far do embryos accumulate material
% do embryos migrate at all? might need to run a simulation without gas to demonstrate whether this is due to PDM
% major vs minor mergers?
